Advancing Sustainable Mobility: A GIS-Based Multi-Criteria Decision Analysis for Equitable Electric Vehicle Supply Equipment Deployment in Philadelphia

### 2025-02-17 code/MCDA-EVSE.Rmd
wsm local mcda

@@

# Get the original rank and compute percentile rank 
# then join it with the result from local MCDA
# to compare whether the percentile rank is higher or lower than the global MCDA
final_percentile <- final %>% 
  dplyr::select(NAME, uniqueID, avg_rank, topsis_rank, wsm_rank) %>%
  mutate(
    avg_percentile = percent_rank(avg_rank) * 100,
    topsis_percentile = percent_rank(topsis_rank) * 100)

@@

# an overall regression summary table here

>>>>

## Try Local WSM

```{r}
# weights_c1 <- c(0.12, 0.14, 0.12, 0.16, 0.16, 0.08, 0.06, 0.06, 0.10)
WSM_c1 <- cluster1 %>% 
  mutate(scl_density = scale_values(density),
         scl_acc = scale_values(TSFCA),
         scl_drivingpop = scale_values(driving_pop),
         scl_evse = scale_values(evse.nn),
         scl_elec = scale_values(elec_cnt),
         scl_parking = scale_values(parking.nn),
         scl_policefire = scale_values(policefire.nn),
         scl_elect_use = scale_values(elect_use)
         ) %>% 
  mutate(scl_evse_re = 0 - scl_evse + 1,
         scl_acc_re = 0 - scl_acc + 1, 
         scl_parking_re = 0 - scl_parking + 1,
         scl_policefire_re = 0 - scl_policefire + 1) %>% 
  mutate(score = 0.12*scl_evse_re + 0.14*scl_density + 0.12*scl_drivingpop + 0.16*scl_acc_re + 0.16*scl_elec + 0.08*scl_parking_re + 0.06*scl_policefire_re + 0.06*zone_score + 0.10*scl_elect_use) %>% 
  dplyr::select(c(uniqueID, score)) %>% 
  mutate(wsm_rank_c1 = rank(score)) %>% 
  mutate(wsm_percentile_c1 = percent_rank(wsm_rank_c1) * 100)

#weights_c2 <- c(0.14, 0.08, 0.08, 0.10, 0.12, 0.12, 0.06, 0.10, 0.20)
WSM_c2 <- cluster2 %>% 
  mutate(scl_density = scale_values(density),
         scl_acc = scale_values(TSFCA),
         scl_drivingpop = scale_values(driving_pop),
         scl_evse = scale_values(evse.nn),
         scl_elec = scale_values(elec_cnt),
         scl_parking = scale_values(parking.nn),
         scl_policefire = scale_values(policefire.nn),
         scl_elect_use = scale_values(elect_use)
         ) %>% # now inverse some scales
  mutate(scl_evse_re = 0 - scl_evse + 1,
         scl_acc_re = 0 - scl_acc + 1, 
         scl_parking_re = 0 - scl_parking + 1,
         scl_policefire_re = 0 - scl_policefire + 1) %>% 
  mutate(score = 0.14*scl_evse_re + 0.08*scl_density + 0.08*scl_drivingpop + 0.10*scl_acc_re + 0.12*scl_elec + 0.12*scl_parking_re + 0.06*scl_policefire_re + 0.10*zone_score + 0.20*scl_elect_use) %>% 
  dplyr::select(c(uniqueID, score)) %>%
  mutate(wsm_rank_c2 = rank(score)) %>%
  mutate(wsm_percentile_c2 = percent_rank(wsm_rank_c2) * 100)

#weights_c3 <- c(0.20, 0.10, 0.12, 0.14, 0.18, 0.06, 0.06, 0.06, 0.08)
WSM_c3 <- cluster3 %>% 
  mutate(scl_density = scale_values(density),
         scl_acc = scale_values(TSFCA),
         scl_drivingpop = scale_values(driving_pop),
         scl_evse = scale_values(evse.nn),
         scl_elec = scale_values(elec_cnt),
         scl_parking = scale_values(parking.nn),
         scl_policefire = scale_values(policefire.nn),
         scl_elect_use = scale_values(elect_use)
         ) %>% 
  mutate(scl_evse_re = 0 - scl_evse + 1,
         scl_acc_re = 0 - scl_acc + 1, 
         scl_parking_re = 0 - scl_parking + 1,
         scl_policefire_re = 0 - scl_policefire + 1) %>% 
  mutate(score = 0.20*scl_evse_re + 0.10*scl_density + 0.12*scl_drivingpop + 0.14*scl_acc_re + 0.18*scl_elec + 0.06*scl_parking_re + 0.06*scl_policefire_re + 0.06*zone_score + 0.08*scl_elect_use) %>% 
  dplyr::select(c(uniqueID, score)) %>%
  mutate(wsm_rank_c3 = rank(score)) %>%
  mutate(wsm_percentile_c3 = percent_rank(wsm_rank_c3) * 100)

#weights_c4 <- c(0.16, 0.14, 0.12, 0.18, 0.12, 0.08, 0.04, 0.06, 0.10)
WSM_c4 <- cluster4 %>% 
  mutate(scl_density = scale_values(density),
         scl_acc = scale_values(TSFCA),
         scl_drivingpop = scale_values(driving_pop),
         scl_evse = scale_values(evse.nn),
         scl_elec = scale_values(elec_cnt),
         scl_parking = scale_values(parking.nn),
         scl_policefire = scale_values(policefire.nn),
         scl_elect_use = scale_values(elect_use)
         ) %>% 
  mutate(scl_evse_re = 0 - scl_evse + 1,
         scl_acc_re = 0 - scl_acc + 1, 
         scl_parking_re = 0 - scl_parking + 1,
         scl_policefire_re = 0 - scl_policefire + 1) %>% 
  mutate(score = 0.16*scl_evse_re + 0.14*scl_density + 0.12*scl_drivingpop + 0.18*scl_acc_re + 0.12*scl_elec + 0.08*scl_parking_re + 0.04*scl_policefire_re + 0.06*zone_score + 0.10*scl_elect_use) %>% 
  dplyr::select(c(uniqueID, score)) %>%
  mutate(wsm_rank_c4 = rank(score)) %>%
  mutate(wsm_percentile_c4 = percent_rank(wsm_rank_c4) * 100)

```{r test with original}

# using original weights
# this is evse.nn, density, driving_pop, TSFCA, elec_cnt, parking.nn, policefire.nn, zone score, elect_use
# weights_original <- c(0.18, 0.12, 0.1, 0.12, 0.15, 0.1, 0.05, 0.05, 0.13)

WSM_c1_org <- cluster1 %>% 
  mutate(scl_density = scale_values(density),
         scl_acc = scale_values(TSFCA),
         scl_drivingpop = scale_values(driving_pop),
         scl_evse = scale_values(evse.nn),
         scl_elec = scale_values(elec_cnt),
         scl_parking = scale_values(parking.nn),
         scl_policefire = scale_values(policefire.nn),
         scl_elect_use = scale_values(elect_use)
         ) %>% 
  mutate(scl_evse_re = 0 - scl_evse + 1,
         scl_acc_re = 0 - scl_acc + 1, 
         scl_parking_re = 0 - scl_parking + 1,
         scl_policefire_re = 0 - scl_policefire + 1) %>% 
  mutate(score = 0.18*scl_evse_re + 0.12*scl_density + 0.1*scl_drivingpop + 0.12*scl_acc_re + 0.15*scl_elec + 0.1*scl_parking_re + 0.05*scl_policefire_re + 0.05*zone_score + 0.13*scl_elect_use) %>% 
  dplyr::select(c(uniqueID, score)) %>% 
  mutate(wsm_rank_c1_org = rank(score)) %>% 
  mutate(wsm_percentile_c1_org = percent_rank(wsm_rank_c1_org) * 100)

WSM_c2_org <- cluster2 %>%
  mutate(scl_density = scale_values(density),
         scl_acc = scale_values(TSFCA),
         scl_drivingpop = scale_values(driving_pop),
         scl_evse = scale_values(evse.nn),
         scl_elec = scale_values(elec_cnt),
         scl_parking = scale_values(parking.nn),
         scl_policefire = scale_values(policefire.nn),
         scl_elect_use = scale_values(elect_use)
         ) %>% 
  mutate(scl_evse_re = 0 - scl_evse + 1,
         scl_acc_re = 0 - scl_acc + 1, 
         scl_parking_re = 0 - scl_parking + 1,
         scl_policefire_re = 0 - scl_policefire + 1) %>% 
  mutate(score = 0.18*scl_evse_re + 0.12*scl_density + 0.1*scl_drivingpop + 0.12*scl_acc_re + 0.15*scl_elec + 0.1*scl_parking_re + 0.05*scl_policefire_re + 0.05*zone_score + 0.13*scl_elect_use) %>% 
  dplyr::select(c(uniqueID, score)) %>%
  mutate(wsm_rank_c2_org = rank(score)) %>%
  mutate(wsm_percentile_c2_org = percent_rank(wsm_rank_c2_org) * 100)

WSM_c3_org <- cluster3 %>%
  mutate(scl_density = scale_values(density),
         scl_acc = scale_values(TSFCA),
         scl_drivingpop = scale_values(driving_pop),
         scl_evse = scale_values(evse.nn),
         scl_elec = scale_values(elec_cnt),
         scl_parking = scale_values(parking.nn),
         scl_policefire = scale_values(policefire.nn),
         scl_elect_use = scale_values(elect_use)
         ) %>% 
  mutate(scl_evse_re = 0 - scl_evse + 1,
         scl_acc_re = 0 - scl_acc + 1, 
         scl_parking_re = 0 - scl_parking + 1,
         scl_policefire_re = 0 - scl_policefire + 1) %>% 
  mutate(score = 0.18*scl_evse_re + 0.12*scl_density + 0.1*scl_drivingpop + 0.12*scl_acc_re + 0.15*scl_elec + 0.1*scl_parking_re + 0.05*scl_policefire_re + 0.05*zone_score + 0.13*scl_elect_use) %>% 
  dplyr::select(c(uniqueID, score)) %>%
  mutate(wsm_rank_c3_org = rank(score)) %>%
  mutate(wsm_percentile_c3_org = percent_rank(wsm_rank_c3_org) * 100)

WSM_c4_org <- cluster4 %>%
  mutate(scl_density = scale_values(density),
         scl_acc = scale_values(TSFCA),
         scl_drivingpop = scale_values(driving_pop),
         scl_evse = scale_values(evse.nn),
         scl_elec = scale_values(elec_cnt),
         scl_parking = scale_values(parking.nn),
         scl_policefire = scale_values(policefire.nn),
         scl_elect_use = scale_values(elect_use)
         ) %>% 
  mutate(scl_evse_re = 0 - scl_evse + 1,
         scl_acc_re = 0 - scl_acc + 1, 
         scl_parking_re = 0 - scl_parking + 1,
         scl_policefire_re = 0 - scl_policefire + 1) %>% 
  mutate(score = 0.18*scl_evse_re + 0.12*scl_density + 0.1*scl_drivingpop + 0.12*scl_acc_re + 0.15*scl_elec + 0.1*scl_parking_re + 0.05*scl_policefire_re + 0.05*zone_score + 0.13*scl_elect_use) %>% 
  dplyr::select(c(uniqueID, score)) %>%
  mutate(wsm_rank_c4_org = rank(score)) %>%
  mutate(wsm_percentile_c4_org = percent_rank(wsm_rank_c4_org) * 100)

# Get the original rank and compute percentile rank 
# then join it with the result from local MCDA
# to compare whether the percentile rank is higher or lower than the global MCDA
final_percentile <- final %>% 
  dplyr::select(NAME, uniqueID, avg_rank, topsis_rank, wsm_rank) %>%
  mutate(
    avg_percentile = percent_rank(avg_rank) * 100,
    wsm_percentile = percent_rank(wsm_rank) * 100,
    topsis_percentile = percent_rank(topsis_rank) * 100)

# join with original percentile
WSM_c1 <- WSM_c1 %>% 
  st_drop_geometry() %>%
  left_join(WSM_c1_org %>% dplyr::select(uniqueID, wsm_percentile_c1_org), by = "uniqueID") %>%
  left_join(final_percentile %>% dplyr::select(uniqueID, 
                                               NAME, 
                                               wsm_percentile, 
                                               avg_percentile,
                                          ), by = "uniqueID") %>%
  dplyr::select(NAME, uniqueID, avg_percentile, wsm_percentile, wsm_percentile_c1, wsm_percentile_c1_org)

WSM_c2 <- WSM_c2 %>%
  st_drop_geometry() %>%
  left_join(WSM_c2_org %>% dplyr::select(uniqueID, wsm_percentile_c2_org), by = "uniqueID") %>%
  left_join(final_percentile %>% dplyr::select(uniqueID, 
                                               NAME, 
                                               wsm_percentile, 
                                               avg_percentile,

                                          ), by = "uniqueID") %>%
  dplyr::select(NAME, uniqueID, avg_percentile, wsm_percentile, wsm_percentile_c2, wsm_percentile_c2_org)

WSM_c3 <- WSM_c3 %>%
  st_drop_geometry() %>%
  left_join(WSM_c3_org %>% dplyr::select(uniqueID, wsm_percentile_c3_org), by = "uniqueID") %>%
  left_join(final_percentile %>% dplyr::select(uniqueID, 
                                               NAME, 
                                               wsm_percentile, 
                                               avg_percentile,

                                          ), by = "uniqueID") %>%
  dplyr::select(NAME, uniqueID, avg_percentile, wsm_percentile, wsm_percentile_c3, wsm_percentile_c3_org)

WSM_c4 <- WSM_c4 %>%
  st_drop_geometry() %>%
  left_join(WSM_c4_org %>% dplyr::select(uniqueID, wsm_percentile_c4_org), by = "uniqueID") %>%
  left_join(final_percentile %>% dplyr::select(uniqueID, 
                                               NAME, 
                                               wsm_percentile, 
                                               avg_percentile,

                                          ), by = "uniqueID") %>%
  dplyr::select(NAME, uniqueID, avg_percentile, wsm_percentile, wsm_percentile_c4, wsm_percentile_c4_org)

WSM_c1 %>% 
  ggplot(aes(x = wsm_percentile , y = wsm_percentile_c1)) +
  geom_point() +
  geom_smooth(method = "lm", color = "blue", se = FALSE) + 
  labs(title = "Comparison of Global and Local TOPSIS Percentile Rank for Cluster 1",
       x = "Global Percentile Rank",
       y = "Local Percentile Rank") +
  theme_minimal() +
  theme(plot.subtitle = element_text(size = 12,face = "italic"),
        plot.title = element_text(size = 15, face = "bold"), 
        axis.text.x=element_text(size=8),
        axis.text.y=element_text(size=8), 
        axis.title=element_text(size=9), 
        panel.background = element_blank(),
        panel.border = element_rect(colour = "grey", fill=NA, linewidth =0.8))

model1_wsm <- lm(wsm_percentile_c4 ~ wsm_percentile, data = WSM_c4)
model2_wsm<- lm(wsm_percentile_c4 ~ wsm_percentile_c4_org, data = WSM_c4)
model3_wsm <- lm(wsm_percentile ~ wsm_percentile_c4_org, data = WSM_c4)

library(stargazer)
stargazer(model1_wsm, model2_wsm, model3_wsm, type = "text")
```

## Try Local TOPSIS

@@

@@
